### 2013-08-02 Docs/MESA/instructions.tex
latest MESA docs

To install {\sf MESA}, download the source code from the {\sf MESA} website, 
{\tt http://mesa.sourceforge.net}. Check the website for the latest version, 
which is 4849 when the last upgrades to this document were added. The codes 
and algorithms 
discussed here were created when the current version of {\sf MESA} was 
version 4088 and have been modified accordingly to keep up with the quickly 
channging source code. It might be possible to use older versions of 
{\sf MESA} with 
the information in this document, although there are minor changes and 
upgrades between {\sf MESA} versions that may cause the routines and algorithms 
discussed here to crash. To download the source code associated with version 
4849, issue the following command:
\begin{lstlisting}[language=bash,mathescape=false]
  svn co -r 4849 svn://svn.code.sf.net/p/mesa/code/trunk mesa
\end{lstlisting}
This will download the source into a directory called {\tt mesa} located 
in the directory where this command was run. This step might take some time 
($\sim$3 min depending on the internet connection) possibly because the 
final {\tt mesa} directory will take up roughly 2.1 GB.

Rich Townsend has developed a Software Development Kit (SDK) to help install 
{\sf MESA} called {\sf MESASDK}. The SDK will install certain software 
packages that {\sf MESA} uses and it will also change the {\tt PATH} 
environment variable such that these programs become the default. A tarball 
of the SDK software and indepth instructions for its use are available on 
Townsend's website:
\begin{lstlisting}[language=html,mathescape=false]
  http://www.astro.wisc.edu/~townsend/static.php?ref=mesasdk
\end{lstlisting}
The instructions in this document do not use the SDK in order to avoid 
automatically changing the {\tt PATH} variable.

@@

%-------------------------------------------------------------------------
\subsubsection{Environment Variables (NOT using SDK)}
A few environment variables need to be set (commands are shown for the 
Bourne shell):
\begin{enumerate}
\item Set location of {\tt mesa}: 
{\tt export MESA\_DIR=<absolute-path-to-mesa-directory>}

@@

%-------------------------------------------------------------------------
\subsubsection{Modify the Makefile (NOT using SDK)}
The main makefile that contains compiler information including flags is 
located in the {\tt mesa/utils} directory and is called 
{\tt makefile\_header\_non\_mesasdk}.

All of the following changes apply to 
variables that already exist. There is no need to explicitly add new code 
unless you intend to compile {\sf MESA} with something other than 
{\tt gfortran} or {\tt ifort}. If you want to compile using something 
different, then the only code that needs to be explicitly added are the 
definitions of the compiler flags.

Open the {\tt makefile\_header\_non\_mesasdk} and change the following:
\begin{enumerate}
\item Compilers: First change the compilers labeled {\tt CC} (C compiler) 
and {\tt FC} (Fortran compiler) to the compilers of your choice ({\sf MESA} 
defaults are {\tt gcc} and {\tt ifort}).

\item Remove the second instance of {\tt CC} located directly under the 
{\tt SPECIAL\_FC\_FLAGS} variable

@@

\item {\tt BLAS} and {\tt LAPACK}: Set the location of {\tt LAPACK} and 
{\tt BLAS} provided with the {\sf MESA} source tree using the following 
definitions (these should be the default settings):
\begin{lstlisting}[language=make,mathescape=false,upquote=true]
  WHICH_LAPACK = USE_SRCS
  LOAD_LAPACK = -lmesalapack
  WHICH_BLAS = USE_SRCS
  LOAD_BLAS = -lmesablas
  MKL_INCLUDE =
\end{lstlisting}

@@

\item {\tt USE\_SE}: Set the variable {\tt USE\_SE} to {\tt NO} and set both 
{\tt LOAD\_SE} and {\tt INCLUDE\_SE} to be empty. These also need to be set 
in the correct if-construct according to which compiler was chosen. 
{\tt SE} is a library of 
formats used for writing/reading data in stellar evolution codes.

@@

%============================================================================
% This next \item is necessary when compiling in serial
%============================================================================
%
\item Edit 1 source file: An undefined reference to 
{\tt omp\_get\_max\_threads} occurs in the {\tt const\_def.f} file located in 
{\tt const/public/const\_def.f} when compiling in serial.

@@

All {\sf MESA} modules have a makefile that includes the main makefile named 
{\tt makefile\_header}. The file that was just edited was named 
{\tt makefile\_header\_non\_mesasdk} and so the name must be changed in order 
for the edits to take effect. This can be done as follows:
\begin{lstlisting}[language=bash,mathescape=false]
  cp makefile_header_non_mesasdk makefile_header
\end{lstlisting}

@@

In the {\tt mesa/utils} directory, open the {\tt makefile\_header} file. 
Change the {\tt FCopenmp} flag ({\tt gfortran} default = {\tt -fopenmp}) to 
the appropriate value so as to disable OpenMP, e.g. with {\tt gfortran} 
{\tt FCopenmp} should be empty.

Anytime a change is made to a makefile or any source code, the module must 
be recompiled, which is a three step process involving the {\tt clean}, 
{\tt mk} and {\tt export} scripts provided in each module directory. The 
first step is to clean out any old files using the {\tt ./clean} script. 
Step two is to make the module using the {\tt ./mk} script. The last step is 
to export the {\tt .mod} files and the library to the right directories using 
the {\tt ./export} script. Section \ref{sec:compile_scripts} discusses a few 
simple scripts that are meant to help with compiling modules.

@@

The {\sf MESA} libraries have the name {\tt lib(module-name).a}, e.g. 
{\tt libeos.a} or {\tt libnet.a}. The libraries need to be linked using the 
{\tt -L} flag with the {\tt mesa/lib} directory and a list of which libraries 
to load, e.g. {\tt -L\$(MESA\_DIR)/lib -leos -lnet -lnum}. {\sf MESA} 
{\tt .mod} files are linked to the code using the {\tt -I} flag and the 
{\tt mesa/include} directory, e.g. {\tt -I\$(MESA\_DIR)/include}.

@@

The wrapper routines that allow {\sf MAESTRO} to use {\sf MESA} modules are 
located in the {\tt AstroDev/EOS/MESA} and {\tt AstroDev/networks/MESA} 
directories. In each directory there is a {\tt GPackage.mak} file that 
specifies the proper {\tt -I} and {\tt -L} compiler flags along with what 
source files to include in the build. To use the EOS, open the 
{\tt GNUmakefile} in the {\sf MAESTRO} problem directory and add the 
following lines:
\begin{lstlisting}[language=make,mathescape=false]
  EOS_TOP_DIR := $(ASTRODEV_DIR)/EOS
  EOS_DIR := MESA
\end{lstlisting}
To use the network, add the following lines:
\begin{lstlisting}[language=make,mathescape=false]
  NETWORK_TOP_DIR := $(ASTRODEV_DIR)/networks
  NETWORK_DIR := MESA
  GENERAL_NET_INPUTS := $(NETWORK_TOP_DIR)/MESA/Net-Files/net_input.net
\end{lstlisting}
The {\tt GENERAL\_NET\_INPUTS} variable holds the path to the
{\tt net\_input.net} file, which does not need to be kept in the
{\tt Net-Files} directory. The {\tt net\_input.net} file should have the 
following format:
\begin{lstlisting}[language=make,mathescape=false]
  # name              short-name    aion   zion
  hydrogen-1             H1          1.0    1.0
  helium-4               He4         4.0    2.0
  carbon-12              C12         12.0   6.0
  oxygen-16              O16         16.0   8.0
  neon-20                Ne20        20.0  10.0
  sodium-23              Na23        23.0  11.0
  magnesium-23           Mg23        23.0  12.0
  magnesium-24           Mg24        24.0  12.0
  silicon-28             Si28        28.0  14.0
\end{lstlisting}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% How to setup/envoke/shutdown MESA modules
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Using {\sf MESA} Modules}

@@

The routines that carry out these processes (setting up EOS, setting up 
isotope map, shutdown routines, etc.) have already been written and live in 
{\sf AstroDev/networks/MESA} and {\sf AstroDev/EOS/MESA}. This section 
outlines the necessary steps taken by those routines in order to set up the 
EOS and network.

@@

\item Chem Initialization: This is done by calling the {\tt chem\_init} 
subroutine: {\tt call chem\_init(data\_dir, `isotopes.data\_real', ierr)}. 
The second argument can also be {\tt `isotopes.data\_approx'}. The ``real'' 
file contains actual values for 
atomic quantities such as atomic weight. The ``approx'' file contains 
approximate values for these quantities, e.g. the mass of the proton and 
neutron are both equal to 1.0. The {\tt ierr} argument is an integer that 
holds the exit status of the subroutine, non-zero means there was an error

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Network
%%%%%%%%%%
\subsection{Network}

@@

There are significantly more steps involved when initializing the {\sf MESA} 
network. Here is a brief outline of what needs to happen:
\begin{enumerate}
\item Get the {\sf MESA} and {\sf AstroDev} directory locations
\item Convert the short species names to {\sf MESA} format
\item Generate a file that will tell {\sf MESA} which reactions to add based 
on the isotopes in the current network ({\tt call which\_reactions})
\item Initialize {\sf MESA} {\tt chem}, {\tt weak} and {\tt reaclib} modules
\item Setup the map between {\sf MAESTRO} and {\sf MESA}, see Section 
\ref{sec:mapping} for a more in depth explanation
\item Initialize {\sf MESA} {\tt rates} and {\tt net} modules
\item Allocate a network handle
\item Call the {\tt net\_start\_def}, {\tt read\_net\_file}, {\tt net\_ptr} 
and {\tt net\_finish\_def} subroutines
\item Output a log file containing the species in the current net and which 
reactions were chosen by {\sf MESA} based on those species
\item Set {\tt which\_rates\_choice} and allocate {\tt which\_rates} array
\item Set entire array equal to {\tt which\_rates\_choice} and call 
{\tt net\_set\_which\_rates}
\item Initialize which linear algebra solver to use
\end{enumerate}

In the {\tt Do\_One\_Zone\_Burn.f90} routine, there are two more ``set up'' 
routines: {\tt set\_rattab\_range} and {\tt net\_setup\_tables}. I have not 
found a good way to call these routines only once with the set up of all the 
other routines because the {\tt set\_rattab\_range} routine depends on 
temperature. As a result, both routines are called every time a burn is to 
take place.

@@

Calling the {\sf MESA} network from within the {\sf MAESTRO} wrapper starts 
with calling a modified {\sf MESA} wrapper:
\begin{lstlisting}[language=fortran,mathescape=false]
  call Do_One_Zone_Burn(dens, temp, dt, Xin, burn_ergs, Xout)
\end{lstlisting}
The inputs are density, temperature, how long to burn and starting mass 
fractions. The outputs are the energy generation rate ({\tt burn\_ergs}) 
in units of ergs g$^{-1}$ s$^{-1}$ and the final mass fractions.

To calculate {\tt rho\_omegadot} ($\rho \dot{\omega}$) and {\tt rho\_Hnuc} 
($\rho H_{{\rm nuc}}$) the following code can be used after a call to 
{\tt Do\_One\_Zone\_Burn}:
\begin{lstlisting}[language=fortran,mathescape=false]
  do i=1, nspec
     dX = Xout(i) - Xin(i)
     rho_omegadot(i) = dens * dX / dt
  enddo

  rho_Hnuc = dens * burn_ergs
\end{lstlisting}
There is a support routine called {\tt burn\_solout} that is passed to the 
{\sf MESA} burner. This routine is called after every step of the burn and 
is responsible for calculating the energy generation rate. {\tt burn\_solout} 
is passed as an argument to the {\tt net\_1\_zone\_burn} subroutine (the 
{\sf MESA} burner) and as a result the actual call to the {\tt burn\_solout} 
routine is buried deep within the {\sf MESA} source code.

@@

Some useful scripts have been written in {\tt BASH} to help with compiling 
{\sf MESA} modules. They live in the {\sf AstroDev/networks/MESA/scripts} 
directory. Currently, there are two scripts; one that compiles all modules 
(which unfortunately is version dependant), and one that compiles a single 
module.

>>>>

To install {\sf MESA}, download the source code from the {\sf MESA} website, 
{\tt http://mesa.sourceforge.net}. Check the website for the latest version, 
which is 5271 when the last upgrades to this document were added. The codes 
and algorithms 
discussed here were created when the current version of {\sf MESA} was 
version 4088 and have been modified accordingly to keep up with the quickly 
channging source code. It might be possible to use older versions of 
{\sf MESA} with 
the information in this document, although there are minor changes and 
upgrades between {\sf MESA} versions that may cause the routines and algorithms 
discussed here to crash. These notes are aimed at working with 
version 4942, to download the source code associated with this version, 
issue the following command:
\begin{lstlisting}[language=bash,mathescape=false]
  svn co -r 4942 svn://svn.code.sf.net/p/mesa/code/trunk mesa
\end{lstlisting}
This will download the source into a directory called {\tt mesa} located 
in the directory where this command was run. This step might take some time 
($\sim$1-3 min depending on the internet connection) possibly because the 
final {\tt mesa} directory will take up roughly 1.8 GB.

Rich Townsend has developed a Software Development Kit (SDK) to help install 
{\sf MESA} called {\sf MESASDK}. The SDK will install certain software 
packages that {\sf MESA} uses and it will also change the {\tt PATH} 
environment variable such that these programs become the default. A tarball 
of the SDK software and indepth instructions for its use are available on 
Townsend's website:
\begin{lstlisting}[language=html,mathescape=false]
  http://www.astro.wisc.edu/~townsend/static.php?ref=mesasdk
\end{lstlisting}
The instructions in this document {\bf \emph{do not}} use the SDK in order 
to avoid automatically changing the {\tt PATH} variable.

A quick outline for compiling {\sf MESA} for use with {\sf MAESTRO} is 
shown below. Detailed instructions are discussed in Section \ref{sec:woSDK}.
\begin{enumerate}
\item Set environment variables, {\tt MESA\_DIR} and {\tt OMP\_NUM\_THREADS}
\item Change name of makefile
\item Makefile: choose compiler, comment out second instance of 
{\tt CC = gcc}, turn off {\tt USE\_PGSTAR} and {\tt LOAD\_PGPLOT}
\item Modify the {\tt const/public/const\_def.f} to fix the 
{\tt omp\_get\_max\_threads()} line
\item Run {\tt install} script
\item Turn off OMP
\item Recompile using the {\tt compile-all-mods.sh} script
\end{enumerate}

@@

%-------------------------------------------------------------------------
\subsubsection{Environment Variables (NOT using SDK)}
A few environment variables need to be set (commands are shown for the 
BASH shell):
\begin{enumerate}
\item Set location of {\tt mesa}: 
{\tt export MESA\_DIR=<absolute-path-to-mesa-directory>}

@@

%-------------------------------------------------------------------------
\subsubsection{Modify the Makefile (NOT using SDK)}
The main makefile that contains compiler information including flags is 
located in the {\tt mesa/utils} directory and is called 
{\tt makefile\_header}. All {\sf MESA} modules have a makefile that 
includes this main makefile. The file that needs to be edited is named 
{\tt makefile\_header\_non\_mesasdk} and so the name must be changed in order 
for the edits to take effect. This can be done as follows:
\begin{lstlisting}[language=bash,mathescape=false]
  cp makefile_header_non_mesasdk makefile_header
\end{lstlisting}

All of the changes listed below apply to variables that already exist. 
There is no need to explicitly add new code 
unless you intend to compile {\sf MESA} with something other than 
{\tt gfortran} or {\tt ifort}. If you want to compile using something 
different, then the only code that needs to be explicitly added are the 
definitions of the compiler flags.

Open the {\tt makefile\_header} file and change the following:
\begin{enumerate}
\item Compilers: First change the compilers labeled {\tt CC} (C compiler) 
and {\tt FC} (Fortran compiler) to the compilers of your choice ({\sf MESA} 
defaults are {\tt gcc} and {\tt ifort}).

\item Remove (or comment out) the second instance of {\tt CC} located 
directly under the {\tt SPECIAL\_FC\_FLAGS} variable

@@

\item {\tt BLAS} and {\tt LAPACK}: Make sure the location of {\tt LAPACK} and 
{\tt BLAS} (provided with the {\sf MESA} source tree) are set using the 
following definitions (these should be the default settings):
\begin{lstlisting}[language=make,mathescape=false,upquote=true]
  WHICH_LAPACK = USE_SRCS
  LOAD_LAPACK = -lmesalapack
  WHICH_BLAS = USE_SRCS
  LOAD_BLAS = -lmesablas
  MKL_INCLUDE =
\end{lstlisting}

@@

\item {\tt USE\_SE}: Make sure the variable {\tt USE\_SE} is set to {\tt NO} 
and both {\tt LOAD\_SE} and {\tt INCLUDE\_SE} are empty. These also need to 
be set in the correct if-construct according to which compiler was chosen. 
{\tt SE} is a library of 
formats used for writing/reading data in stellar evolution codes.

@@

%============================================================================
% This next \item is necessary when compiling in serial
%============================================================================
%
\item Edit 1 source file: An undefined reference to 
{\tt omp\_get\_max\_threads} occurs in the {\tt const\_def.f} file located in 
{\tt const/public/const\_def.f} when running the code in serial.

@@

@@

In the {\tt mesa/utils} directory, open the {\tt makefile\_header} file. 
Change the {\tt FCopenmp} flag ({\tt gfortran} default = {\tt -fopenmp}) to 
the appropriate value so as to disable OpenMP, e.g.\ with {\tt gfortran} 
{\tt FCopenmp} should be empty. Now all modules must be recompiled. This is 
most easily accomplished with the scripts described in Section 
\ref{sec:compile_scripts}.

Anytime a change is made to a makefile or any source code, the module must 
be recompiled, which is a three step process involving the {\tt clean}, 
{\tt mk} and {\tt export} scripts provided in each module directory. The 
first step is to clean out any old files using the {\tt ./clean} script. 
Step two is to make the module using the {\tt ./mk} script. The last step is 
to export the {\tt .mod} files and the library to the right directories using 
the {\tt ./export} script. Section \ref{sec:compile_scripts} discusses a few 
simple scripts that are meant to help with compiling the modules.

@@

The {\sf MESA} libraries have the name {\tt lib(module-name).a}, e.g.\ 
{\tt libeos.a} or {\tt libnet.a}. The libraries need to be linked using the 
{\tt -L} flag with the {\tt mesa/lib} directory and a list of which libraries 
to load, e.g.\ {\tt -L\$(MESA\_DIR)/lib -leos -lnet -lnum}. {\sf MESA} 
{\tt .mod} files are linked to the code using the {\tt -I} flag and the 
{\tt mesa/include} directory, e.g.\ {\tt -I\$(MESA\_DIR)/include}.

@@

The wrapper routines that allow {\sf MAESTRO} to use {\sf MESA} modules are 
located in the {\tt AstroDev/EOS/MESA}, {\tt AstroDev/networks/MESA} and 
{\tt AstroDev/networks/MESA\_SDC} 
directories. In each directory there is a {\tt GPackage.mak} file that 
specifies the proper {\tt -I} and {\tt -L} compiler flags along with what 
source files to include in the build. To use the EOS, open the 
{\tt GNUmakefile} in the {\sf MAESTRO} problem directory and add the 
following lines:
\begin{lstlisting}[language=make,mathescape=false]
  EOS_TOP_DIR := $(ASTRODEV_DIR)/EOS
  EOS_DIR := MESA
\end{lstlisting}
To use the Strang-Split version of the network, add the following lines:
\begin{lstlisting}[language=make,mathescape=false]
  NETWORK_TOP_DIR := $(ASTRODEV_DIR)/networks
  NETWORK_DIR := MESA
  GENERAL_NET_INPUTS := $(NETWORK_TOP_DIR)/$(NETWORK_DIR)/Net-Files/net_input.net
\end{lstlisting}
To use the Spectral Deferred Corrections (SDC) version of the network add 
these lines instead:
\begin{lstlisting}[language=make,mathescape=false]
  SDC := t

  NETWORK_TOP_DIR := $(ASTRODEV_DIR)/networks
  NETWORK_DIR := MESA_SDC
  GENERAL_NET_INPUTS := $(NETWORK_TOP_DIR)/$(NETWORK_DIR)/Net-Files/net_input.net
\end{lstlisting}
The {\tt GENERAL\_NET\_INPUTS} variable holds the path to the
{\tt net\_input.net} file, which does not need to be kept in the
{\tt Net-Files} directory. The {\tt net\_input.net} file should have the 
following format:
\begin{lstlisting}[language=make,mathescape=false]
  # name              short-name    aion   zion
  hydrogen-1             H1          1.0    1.0
  helium-4               He4         4.0    2.0
  carbon-12              C12         12.0   6.0
  oxygen-16              O16         16.0   8.0
  neon-20                Ne20        20.0  10.0
  sodium-23              Na23        23.0  11.0
  magnesium-23           Mg23        23.0  12.0
  magnesium-24           Mg24        24.0  12.0
  silicon-28             Si28        28.0  14.0
\end{lstlisting}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% How to setup/envoke/shutdown MESA modules
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Using {\sf MESA} Modules}
\label{sec:using_modules}

@@

The routines that carry out these processes (setting up EOS, setting up 
isotope map, shutdown routines, etc.) {\bf \emph{have already been written}} 
and live in {\tt AstroDev/networks/MESA}, {\tt AstroDev/networks/MESA\_SDC} 
and {\tt AstroDev/EOS/MESA}. This section 
outlines the necessary steps taken by those routines in order to set up the 
EOS and network.

@@

\item Chem Initialization: This is done by calling the {\tt chem\_init} 
subroutine: {\tt call chem\_init(data\_dir, `isotopes.data\_real', ierr)}. 
The second argument can also be {\tt `isotopes.data\_approx'}. The ``real'' 
file contains actual values for 
atomic quantities such as atomic weight. The ``approx'' file contains 
approximate values for these quantities, e.g.\ the mass of the proton and 
neutron are both equal to 1.0. The {\tt ierr} argument is an integer that 
holds the exit status of the subroutine, non-zero means there was an error

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Network
%%%%%%%%%%
\subsection{Network}
There are currently two different algorithms that integrate the nuclear 
reaction network in {\sf MAESTRO}; Strang Splitting and Spectral Deferred 
Corrections (SDC). The routines to go from {\sf MAESTRO} 
to {\sf MESA} using Strang are stored in {\tt AstroDev/networks/MESA} while 
the routines that implement the SDC method are stored in 
{\tt AstroDev/networks/MESA\_SDC}.

The setup/shutdown procedures for both methods are very similar. The largest 
difference is that the Strang method uses {\sf MESA} to obtain the right 
hand sides of the network equations \emph{and} do the integration, while the 
SDC method uses {\sf MESA} to obtain the right hand sides and uses 
{\sf MAESTRO} to integrate the equations using {\tt VODE}.

@@

There are significantly more steps involved when initializing the {\sf MESA} 
network. Here is a brief outline of what happens:
\begin{enumerate}
\item Get the {\sf MESA} and {\sf AstroDev} directory locations
\item Convert the short species names to {\sf MESA} format
\item Generate a file that will tell {\sf MESA} which reactions to add based 
on the isotopes in the current network
\item Initialize various {\sf MESA} modules
\item Setup the map between {\sf MAESTRO} and {\sf MESA} (see Section 
\ref{sec:mapping} for a more in depth explanation)
\item Initialize {\sf MESA} {\tt rates} and {\tt net} modules
\item Allocate a network handle
\item Call the {\tt net\_start\_def}, {\tt read\_net\_file}, {\tt net\_ptr} 
and {\tt net\_finish\_def} subroutines
\item Output a log file containing the species in the current net and which 
reactions were chosen by {\sf MESA} based on those species
\item Set {\tt which\_rates\_choice} and allocate {\tt which\_rates} array
\item Set entire array equal to {\tt which\_rates\_choice} and call 
{\tt net\_set\_which\_rates}
\item Call {\tt net\_setup\_tables}
\item Set which linear algebra solver to use
\item Set the ODE solver to use for the network integration (Strang-Split 
version only)
\end{enumerate}

@@

In the case of Strang Splitting the call to the {\sf MESA} burner looks like 
this:
\begin{lstlisting}[language=fortran,mathescape=false]
  call net_1_zone_burn( &
        handle, solver_choice, species, num_reactions, 0.0_dp_t, burn_tend, &
        xin, clip, num_times, times, log10Ts_f1, log10Rhos_f1, etas_f1, &
        dxdt_source_term, rate_factors, category_factors, std_reaction_Qs, &
        std_reaction_neuQs, screening_mode, theta_e_for_graboske_et_al, &
        h, max_step_size, max_steps, rtol, atol, itol, &
        decsol_choice, caller_id, burn_solout, iout, ending_x, &
        nfcn, njac, nstep, naccpt, nrejct, time_doing_net, ierr)
\end{lstlisting}
The inputs are density, temperature, how long to burn, starting mass 
fractions, various rates and parameters as well as some things to control the 
integration. The outputs are the energy ({\tt burn\_ergs}) in units of ergs/g, 
the ending mass fractions and various counters such as number of jacobian 
evaluations. This routine will calculate the right hand sides, pass them 
to a stiff ODE integrator and return the final results.

In the case of the SDC method, the call to the {\sf MESA} routine that returns 
the instantaneous right hand sides looks like this:
\begin{lstlisting}[language=fortran,mathescape=false]
  call burn_derivs(nspec, t, xmass(1:nspec), ydot(1:nspec), lrpar, rpar, &
                   lipar, ipar, ierr)
\end{lstlisting}
The inputs are mass fraction, work arrays and the time. The output is the 
time derivative of the mass fraction. This routine is only responsible for 
calculating the right hand sides of the network equations, {\tt VODE} is tasked 
with the integration in this algorithm.

@@

Some useful scripts have been written in {\tt BASH} to help with compiling 
{\sf MESA} modules. They live in the {\tt AstroDev/networks/MESA/scripts} 
directory. Currently, there are two scripts; one that compiles all modules 
(which unfortunately is version dependant), and one that compiles a single 
module.
